---
title: "Outlier Report"
output:
  html_document:
    toc: true
    toc_depth: 3
    theme: null
    df_print: paged
params:
  dataset_name: "dataset"
  dt_path: "data.csv"
  outliers_path: "outliers.csv"
  diagnostics_path: "diagnostics.json"
---

<!-- Theme styles -->
<link rel="stylesheet" href="../assets/light.css" id="theme-light">
<link rel="stylesheet" href="../assets/dark.css"  id="theme-dark" disabled>

<!-- Theme toggle -->
<button class="btn-toggle" id="themeToggle" title="Toggle theme">🌗 Toggle Theme</button>
<script>
(function(){
  const light = document.getElementById('theme-light');
  const dark  = document.getElementById('theme-dark');
  const btn   = document.getElementById('themeToggle');
  const key   = 'odet-theme';
  function setTheme(mode){
    if(mode === 'dark'){ dark.disabled = false; light.disabled = true; }
    else { dark.disabled = true; light.disabled = false; }
    try{ localStorage.setItem(key, mode); }catch(e){}
  }
  const saved = (function(){ try{ return localStorage.getItem(key); }catch(e){ return null; } })();
  setTheme(saved === 'dark' ? 'dark' : 'light');
  btn.addEventListener('click', function(){
    const now = dark.disabled ? 'dark' : 'light';
    setTheme(now);
  });
})();
</script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)

suppressPackageStartupMessages({
  library(data.table)
  library(jsonlite)
  library(ggplot2)
})

has_dt_pkg <- requireNamespace("DT", quietly = TRUE)
if (has_dt_pkg) { DT <- asNamespace("DT") }

ds_name  <- params$dataset_name
dt       <- tryCatch(fread(params$dt_path), error = function(e) NULL)
outliers <- tryCatch(fread(params$outliers_path), error = function(e) NULL)
diag     <- tryCatch(jsonlite::fromJSON(params$diagnostics_path, simplifyVector = FALSE),
                     error = function(e) NULL)

has_dt   <- !is.null(dt)
has_out  <- !is.null(outliers) && nrow(outliers) > 0
has_diag <- !is.null(diag)

render_table <- function(df, ...) {
  if (has_dt_pkg) DT$datatable(df, options=list(pageLength=10, scrollX=TRUE), rownames=FALSE, ...)
  else df
}
```

# Summary

**Dataset:** `r ds_name`  
**Rows:** `r if (has_dt) nrow(dt) else "?"` • **Columns:** `r if (has_dt) ncol(dt) else "?"`  

**Artifacts:**  
- Raw data: `r params$dt_path`  
- Outliers CSV: `r params$outliers_path`  
- Diagnostics JSON: `r params$diagnostics_path`

---

# Detector Settings & Thresholds

```{r}
if (has_diag) {
  dets <- diag$detectors
  df <- data.frame(detector=names(dets), t(sapply(dets, function(x) unlist(x))),
                   check.names=FALSE, row.names=NULL)
  print(render_table(df))
}
```

---

# Top Outliers

```{r}
if (has_diag && !is.null(diag$top_flags)) {
  for (d in names(diag$top_flags)) {
    cat("## ", d, "\n", sep="")
    tf <- diag$top_flags[[d]]
    if (length(tf)) print(render_table(as.data.frame(tf, check.names=FALSE)))
  }
}
```

---

# Per-Column Diagnostics

```{r}
if (has_diag && length(diag$columns)) {
  for (col in names(diag$columns)) {
    cd <- diag$columns[[col]]
    cat("## ", col, "\n", sep="")
    tab <- data.frame(
      metric = c("n","n_na","pct_na","mean","sd","min","q1","median","q3","max",
                 "IQR_lower","IQR_upper","MAD_lower","MAD_upper"),
      value  = c(cd$counts$n, cd$counts$n_na, cd$counts$pct_na,
                 cd$summary$mean, cd$summary$sd, cd$summary$min,
                 cd$summary$q1, cd$summary$median, cd$summary$q3, cd$summary$max,
                 cd$iqr$lower, cd$iqr$upper, cd$mad$lower, cd$mad$upper),
      row.names=NULL, check.names=FALSE
    )
    print(render_table(tab))

    cat("<details><summary>Visuals</summary>\n")
    if (has_dt && col %in% names(dt)) {
      x <- suppressWarnings(as.numeric(dt[[col]]))
      dfp <- data.frame(x=x)
      print(ggplot(dfp, aes(x=x)) + geom_histogram(bins=30))
      print(ggplot(dfp, aes(y=x, x=1)) + geom_boxplot(outlier.shape=NA))
    }
    cat("</details>\n")
  }
}
```

---

# Time-Series Anomaly Visuals

```{r}
if (has_diag && !is.null(diag$timeseries) && has_dt) {
  time_col <- diag$timeseries$time_col
  if (requireNamespace("anomalize", quietly=TRUE)) {
    library(anomalize); library(dplyr); library(tibble)
    for (col in names(diag$columns)) {
      if (!col %in% names(dt)) next
      tvec <- suppressWarnings(as.Date(dt[[time_col]]))
      x <- suppressWarnings(as.numeric(dt[[col]]))
      ord <- order(tvec); df <- tibble(time=tvec[ord], value=x[ord])
      if (nrow(df)<10) next
      md <- suppressMessages(
        df %>% time_decompose(value, method="stl") %>%
          anomalize(remainder, alpha=0.05) %>%
          time_recompose()
      )
      cat("## Anomalize: ", col, "\n", sep="")
      print(ggplot(md, aes(x=time, y=observed)) +
              geom_line() +
              geom_point(data=subset(md, anomaly=="Yes"), aes(y=observed), color="red"))
    }
  }
}
```

---

# Full Outliers Table

```{r}
if (has_out) print(render_table(outliers))
```
